---
title: "income maps"
author: "Ivanna Rodríguez"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# install.packages(pacman)
pacman::p_load(sf, tigris, tidyverse, tidycensus, readxl, magrittr, here, arrow, patchwork, ggspatial)
```

```{r}
folder_10 <- 'E:/acs5_block_2010_data/'
folder_15 <- 'E:/acs5_block_2015_data/'
folder_20 <- 'E:/acs5_block_2020_data/'

plots_folder <- 'plots/'
```

```{r}
states <- c("IL", "NY", "TX", "GA")
cities <- c("Chicago", "New York", "Houston", "Atlanta")

get_fltr_data <- function(var) {
  df_10 <- read_parquet(paste0(folder_10, 
                             paste0("dkl_", var, "_10_blkgrp_all_states.parquet"))) %>%
  ungroup() %>%
  filter(state_codes %in% states,
         str_detect(cbsa_title, paste(cities, collapse = "|"))) %>%
    select(block_fips, county_fips, county_name, cbsa_fips, cbsa_title,
           state_codes, state_name, group_label, label, block_estimate, block_total,
           county_estimate, county_total, dkl_block, dkl_bin) %>%
    mutate(year = 2010)
  
  df_15 <- read_parquet(paste0(folder_15, 
                               paste0("dkl_", var, "_15_blkgrp_all_states.parquet"))) %>%
    ungroup() %>%
    filter(state_codes %in% states,
         str_detect(cbsa_title, paste(cities, collapse = "|"))) %>%
    select(block_fips, county_fips, county_name, cbsa_fips, cbsa_title,
           state_codes, state_name, group_label, label, block_estimate, block_total,
           county_estimate, county_total, dkl_block, dkl_bin) %>%
    mutate(year = 2015)
  
  df_20 <- read_parquet(paste0(folder_20, 
                               paste0("dkl_", var, "_20_blkgrp_all_states.parquet"))) %>%
    ungroup() %>%
    filter(state_codes %in% states,
         str_detect(cbsa_title, paste(cities, collapse = "|"))) %>%
    select(block_fips, county_fips, county_name, cbsa_fips, cbsa_title,
           state_codes, state_name, group_label, label, block_estimate, block_total,
           county_estimate, county_total, dkl_block, dkl_bin) %>%
    mutate(year = 2020)
  
  return(rbind(df_10, df_15, df_20))
}
```

```{r}
income <- get_fltr_data("income")
```

# Get Shapefiles

Here, I am using the block groups shapefiles from `tidycensus` and filtering to only include block groups within a particular CBSA area. For 2010 and 2015, the shapefile I am using is a 2015 shapefile, whereas for year 2020 I use a 2020 shapefile. This is because the 2020 GEOIDs for census block groups seem to have changed. So the maps for 2010 and 2015 look incomplete when using 2020 shapefiles because the names cannot be matched.


```{r}
state_codes <- income %>% 
  select(block_fips) %>% 
  mutate(state_code = str_sub(block_fips, 1, 2)) %>%
  select(state_code) %>%
  distinct() %>%
  pull(state_code)

counties <- income %>%
  select(county_fips, county_name, cbsa_title) %>%
  mutate(county_fips = str_sub(county_fips, 3, 5)) %>%
  distinct()
```


```{r}
get_block_shp <- function(state_code, county_codes, yr) {
  get_acs(year = yr, 
           geography = "block group", 
           survey = 'acs5', 
           variables = 'B19001_001', 
           state = state_code,
           county = county_codes,
           geometry = TRUE) %>%
    select(GEOID) %>% 
    st_transform(crs = st_crs(4326))
}
```

```{r}
atl_shp_15 <- get_block_shp('13', counties %>% 
                           filter(str_detect(cbsa_title, "Atlanta")) %>% 
                           pull(county_fips), 2015
                         )

atl_shp_20 <- get_block_shp('13', counties %>% 
                           filter(str_detect(cbsa_title, "Atlanta")) %>% 
                           pull(county_fips), 2020
                         )
```
```{r}
nyc_shp_15 <- get_block_shp('36', counties %>% 
                           filter(str_detect(cbsa_title, "New York")) %>% 
                           pull(county_fips), 2015)

nyc_shp_20 <- get_block_shp('36', counties %>% 
                           filter(str_detect(cbsa_title, "New York")) %>% 
                           pull(county_fips), 2020)
```
```{r}
chicago_shp_15 <- get_block_shp('17', counties %>% 
                           filter(str_detect(cbsa_title, "Chicago")) %>% 
                           pull(county_fips), 2015)

chicago_shp_20 <- get_block_shp('17', counties %>% 
                           filter(str_detect(cbsa_title, "Chicago")) %>% 
                           pull(county_fips), 2020)
```

```{r}
houston_shp_15 <- get_block_shp('48', counties %>% 
                           filter(str_detect(cbsa_title, "Houston")) %>% 
                           pull(county_fips), 2015)

houston_shp_20 <- get_block_shp('48', counties %>% 
                           filter(str_detect(cbsa_title, "Houston")) %>% 
                           pull(county_fips), 2020)
```

```{r}
my_theme <- function() {
  theme_minimal() +
    theme(
      plot.title = element_text(size = 16, face = "bold"),
      plot.subtitle = element_text(size = 14),
      axis.title = element_text(size = 12, face = "bold"),
      axis.text = element_text(size = 10),
      legend.title = element_text(size = 10, face = "bold"),
      legend.text = element_text(size = 8),
    )
}
```

# Generate Maps

```{r}
get_spatial_income_data <- function(income, city, yr, shp) {
  df <- income %>%
    filter(str_detect(cbsa_title, city),
           year == yr)
  
  shp %<>%
    left_join(df, by = c("GEOID" = "block_fips")) %>%
    select(GEOID, geometry, dkl_block) %>%
    distinct()
  
  return(shp)
}


format_bin_ranges <- function(vec) {
  # return equally distributed bins for numeric vector column
  breaks <- quantile(vec, probs = seq(0, 1, by = 0.25), na.rm = TRUE)
  labels <- c()
  for (i in 1:(length(breaks) - 1)) {
    # adjust lower bound to avoid repetition
    lower_bound <- ifelse(i == 1, breaks[i], breaks[i] + 0.01)
    
    labels <- c(labels, paste0(format(round(lower_bound, 2), nsmall = 2),
                                "-",
                                format(round(breaks[i+1], 2), nsmall = 2)))
  }
  return(labels)
}


map_dkl_block <- function(income, city, yr, shp) {
  
  shp_map <- get_spatial_income_data(income, city, yr, shp)
  
  bin_labels <- format_bin_ranges(shp_map$dkl_block)
  
  shp_map %<>%
    mutate(dkl_block_group = cut(dkl_block, 
                               breaks = quantile(dkl_block, 
                                                 probs = seq(0, 1, by = 0.25), 
                                                 na.rm = TRUE),
                               labels = bin_labels,
                               include.lowest = TRUE))
  
  cbsa <- df %>% select(cbsa_title) %>% distinct() %>% pull(cbsa_title)
  
  custom_colors <- c("#404040", "#b1b1b1", "#F7C3AB", "#CA0020")
  
  shp_map %>%
    ggplot() +
    annotation_map_tile("cartolight", zoomin = 0) +
    geom_sf(aes(fill = dkl_block_group), color = NA, alpha = 0.6, size = 0.05) +
    scale_fill_manual(values = custom_colors, 
                      breaks = bin_labels,
                      labels = bin_labels,
                      name = "DKL (bits)") +
    labs(title = cbsa,
         subtitle = paste("DKL for Income in", yr)) +
    theme(axis.text = element_blank(),
          axis.title = element_blank(),
          axis.ticks = element_blank())
}
```

```{r}
# format_bin_ranges <- function(vec) {
#   # return equally distributed bins for numeric vector column
#   breaks <- quantile(vec, probs = seq(0, 1, by = 0.25), na.rm = TRUE)
#   labels <- c()
#   for (i in 1:(length(breaks) - 1)) {
#     # adjust lower bound to avoid repetition
#     lower_bound <- ifelse(i == 1, breaks[i], breaks[i] + 0.01)
#     
#     labels <- c(labels, paste0(format(round(lower_bound, 2), nsmall = 2),
#                                 "-",
#                                 format(round(breaks[i+1], 2), nsmall = 2)))
#   }
#   return(labels)
# }
# 
# 
# map_dkl_block <- function(income, city, yr, shp) {
#   df <- income %>%
#     filter(str_detect(cbsa_title, city))
#   
#   # max_range <- df %>%
#   #   group_by(year) %>%
#   #   summarise(max_dkl = max(dkl_block, na.rm = TRUE)) %>%
#   #   pull(max_dkl) %>%
#   #   max()
#   
#   df %<>% filter(year == yr)
#   
#   bin_labels <- format_bin_ranges(shp$dkl_block)
#   
#   shp %<>%
#     left_join(df, by = c("GEOID" = "block_fips")) %>%
#     select(GEOID, geometry, dkl_block) %>%
#     distinct() %>%
#     mutate(dkl_block_group = cut(dkl_block, 
#                                breaks = quantile(dkl_block, 
#                                                  probs = seq(0, 1, by = 0.25), 
#                                                  na.rm = TRUE),
#                                labels = bin_labels,
#                                include.lowest = TRUE))
#   
#   cbsa <- df %>% select(cbsa_title) %>% distinct() %>% pull(cbsa_title)
#   
#   custom_colors <- c("#404040", "#b1b1b1", "#F7C3AB", "#CA0020")
#   
#   shp %>%
#     ggplot() +
#     annotation_map_tile("cartolight", zoomin = 0) +
#     geom_sf(aes(fill = dkl_block_group), color = NA, alpha = 0.6, size = 0.05) +
#     scale_fill_manual(values = custom_colors, 
#                       breaks = bin_labels,
#                       labels = bin_labels,
#                       name = "DKL")
#   
#   # shp %>%
#   #   ggplot() +
#   #   annotation_map_tile("cartolight", zoomin = 0) +
#   #   geom_sf(aes(fill = dkl_block_group), color = NA, alpha = 0.7, size = 0.05) +
#   #   scale_fill_manual(values = custom_colors, breaks = bin_labels, labels = bin_labels,
#   #                     name = "DKL") +
#   #   # scale_fill_viridis_c(option = "magma", direction = -1,
#   #   #                      limits = c(0, max_range), na.value = "lightgray") +
#   #   # scale_fill_gradient(low = "black", high = "darkred", 
#   #   #                     limits = c(0, max_range), na.value = "white") +
#   #   labs(title = cbsa,
#   #        subtitle = paste("DKL for Income in", yr),
#   #        fill = "DKL (bits)") +  # Adding title to the legend
#   #   my_theme() +
#   #   theme(legend.key.width = unit(0.5, "cm"),
#   #         legend.key.height = unit(0.5, "cm"),
#   #         legend.text = element_text(size = 7),
#   #         axis.text = element_blank(),
#   #         axis.title = element_blank())
# }
```


```{r}
atl_10 <- map_dkl_block(income, "Atlanta", "2010", atl_shp_15)
atl_15 <- map_dkl_block(income, "Atlanta", "2015", atl_shp_15)
atl_20 <- map_dkl_block(income, "Atlanta", "2020", atl_shp_20)
```

```{r}
atl_10
ggsave(paste0(plots_folder, "map10_atl.png"), dpi = 300)

atl_15
ggsave(paste0(plots_folder, "map15_atl.png"), dpi = 300)

atl_20
ggsave(paste0(plots_folder, "map20_atl.png"), dpi = 300)
```
```{r}
nyc_10 <- map_dkl_block("New York", "2010")
nyc_15 <- map_dkl_block("New York", "2015")
nyc_20 <- map_dkl_block("New York", "2020")
```

```{r}
nyc_10
ggsave(paste0(plots_folder, "map10_nyc.png"), dpi = 300)

nyc_15
ggsave(paste0(plots_folder, "map15_nyc.png"), dpi = 300)

nyc_20
ggsave(paste0(plots_folder, "map20_nyc.png"), dpi = 300)
```

```{r}
hst_10 <- map_dkl_block("Houston", "2010")
hst_15 <- map_dkl_block("Houston", "2015")
hst_20 <- map_dkl_block("Houston", "2020")
```

```{r}
hst_10
ggsave(paste0(plots_folder, "map10_houston.png"), dpi = 300)

hst_15
ggsave(paste0(plots_folder, "map15_houston.png"), dpi = 300)

hst_20
ggsave(paste0(plots_folder, "map20_houston.png"), dpi = 300)
```

```{r}
chi_10 <- map_dkl_block(income, "Chicago", "2010", chicago_shp_15)
chi_15 <- map_dkl_block(income, "Chicago", "2015", chicago_shp_15)
chi_20 <- map_dkl_block(income, "Chicago", "2020", chicago_shp_20)
```

```{r}
chi_10
ggsave(paste0(plots_folder, "map10_chicago.png"), dpi = 300)

chi_15 
ggsave(paste0(plots_folder, "map15_chicago.png"), dpi = 300)

chi_20
ggsave(paste0(plots_folder, "map20_chicago.png"), dpi = 300)
```

```{r}
knitr::opts_chunk$set(dev = "ragg_png")
```

```{r}
income %>%
  group_by(cbsa_title, year) %>%
  summarise(min_dkl = min(dkl_block),
            max_dkl = max(dkl_block))
```





