---
title: "Chicago Figures"
author: "Ivanna Rodríguez"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

### Read in the data

Here, should probably filter data into their respective categories (i.e. race, income, etc.) so as to have one dataset per each variable so that the plotting function does not have to filter the dataset every time and it runs faster.

The geojsons currently imported all come from tidycensus? but i don't know if they are currently filtered to only include the metro (but probably)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) # , fig.width = 12, fig.height = 12
```

```{r}
# install.packages("pacman")
pacman::p_load(tidyverse, arrow, sf, tidycensus, viridis, patchwork, here, magrittr)
source(here("one-way-dkl/chicago_block/plotting_functions.R"))
```

```{r, include=FALSE}
# dkl/census variable files
df15_dkl <- read_parquet(here("one-way-dkl/data/test_dkl_block_15.parquet"))
df20_dkl <- read_parquet(here("one-way-dkl/data/test_dkl_block_20.parquet"))
all_dkl <- bind_rows(
  mutate(df15_dkl, year = 2015),
  mutate(df20_dkl, year = 2020)
)

# spatial files
geom_block_15 <- st_read(here("one-way-dkl/data/geom_block_15.geojson"))
geom_block_20 <- st_read(here("one-way-dkl/data/geom_block_20.geojson"))
community_areas <- st_read(here("one-way-dkl/data/community_areas.geojson"))

# add spatial info to dkl data sets
map_dkl_df15_all_cbsa <- geom_block_15 %>% 
  right_join(df15_dkl, by=c('GEOID'='block_fips'))

map_dkl_df15 <- geom_block_15 %>% 
  right_join(subset(df15_dkl, cbsa_fips == '16980'), by=c('GEOID'='block_fips'))

map_dkl_df20_all_cbsa <- geom_block_20 %>% 
  right_join(df20_dkl, by=c('GEOID'='block_fips'))

map_dkl_df20 <- geom_block_20 %>% 
  right_join(subset(df20_dkl, cbsa_fips == '16980'), by=c('GEOID'='block_fips'))

# join all into one year(?)  or no ....
```

```{r}
community_data <- data.frame(
  community = c("Downtown", "Hyde Park", "Chatham", "Riverdale", "Little Village", "Humboldt Park", "Lincoln Park", "Albany Park", "Rosemont", "Elk Grove Village"),
  lon = c(-87.623177, -87.5917, -87.6146, -87.6149, -87.7050, -87.7213, -87.6488, 
          -87.7280, -87.87324003415466, -87.97473373657915),
  lat = c(41.881832, 41.7948, 41.7401, 41.6476, 41.8445, 41.8991, 41.9255,
          41.9683, 41.98656849330927, 42.002371095468035)
  )
```

```{r}
# create diff datasets for each var
race_15 <- get_var_df(map_dkl_df15, group_lab = "Race/Ethnicity")
race_20 <- get_var_df(map_dkl_df20, group_lab = "Race/Ethnicity")

income_15 <- get_var_df(map_dkl_df15, group_lab = "Household Income")
income_20 <- get_var_df(map_dkl_df20, group_lab = "Household Income")

educ_15 <- get_var_df(map_dkl_df15, group_lab = "Education attainment")
educ_20 <- get_var_df(map_dkl_df20, group_lab = "Education attainment")

emp_15 <- get_var_df(map_dkl_df15, group_lab = "Employment")
emp_20 <- get_var_df(map_dkl_df20, group_lab = "Employment")
```

```{r}
plot_bin_bar_dodge(all_dkl, group_lab = 'Race/Ethnicity', cbsa = '16980')
plot_bin_bar_dodge(all_dkl, group_lab = 'Household Income', cbsa = '16980')
plot_bin_bar_dodge(all_dkl, group_lab = 'Education attainment', cbsa = '16980')
plot_bin_bar_dodge(all_dkl, group_lab = 'Employment', cbsa = '16980')
```



```{r}
# income maps
max_range <- max(max(race_15$dkl_block), max(race_20$dkl_block))

r15 <- map_dkl(race_15, "Race/Ethnicity", "2015", max_range)
r20 <- map_dkl(race_20, "Race/Ethnicity", "2020", max_range)

r15 + r20
```


```{r}
# income maps
max_range <- max(max(income_15$dkl_block), max(income_20$dkl_block))

i15 <- map_dkl(income_15, "Household Income", "2015", max_range)
i20 <- map_dkl(income_20, "Household Income", "2020", max_range)

i15 + i20
```


```{r}
# education maps
max_range <- max(max(educ_15$dkl_block), max(educ_20$dkl_block))

ed15 <- map_dkl(educ_15, "Education attainment", "2015", max_range)
ed20 <- map_dkl(educ_20, "Education attainment", "2020", max_range)

ed15 + ed20
```


```{r}
# employment maps
max_range <- max(2.5, 2.5)

em15 <- map_dkl(emp_15, "Employment", "2015", max_range)
em20 <- map_dkl(emp_20, "Employment", "2020", max_range)

em15 + em20
```




